Supplemental Material for\\Evaluating `Graphical Perception' with CNNs

### 2018-03-31 PAPER/05_position_length_experiment.tex
Update supplemental.

\noindent\textbf{Perceptual Performance.} In the original position-length experiment, Cleveland and McGill report that types 1-5 were post-ordered by their perceptual difficulty with type 1 being the easiest and type 5 the hardest. We report the average MLAE for each type across our networks (and total 56 runs per type) as follows: 
for type 1 $MLAE= 3.956 $ ($SD= 0.274 $),
for type 2 $MLAE= 3.952 $ ($SD= 0.441 $),
for type 3  $MLAE= 4.349 $ ($SD= 0.367 $),
for type 4 $MLAE= 3.668 $ ($SD= 0.256 $), and 
for type 5 $MLAE= 3.902 $ ($SD= 0.253 $). These distributions yield significance ($F_{4,25}=2.815, p<0.05$) but post hoc comparions show that really only type 3 and type 4 differ ($t_{10}=3.406, p<0.01$). This means that the networks do not prefer a certain type on average which leads us to \textbf{partially accept H2.1} and we do not replicate the same pattern as Cleveland and McGill in their human studies even though our networks' performance is clearly worse than their human baseline.
\\~\\
\noindent\textbf{`Multi' Network Performance.} In Cleveland and McGill's original position-length experiment, humans were asked to judge visualizations of types 1-5. In the bottom of Figure~\ref{fig:figure4_mlae} we average the human performances of types 1-5 to create a `multi' human. Similarly, when training our networks, we first limit each training to one stimuli. However, a `multi' network which learns types 1-5 simultaenously is closer to the human in the original user study.

>>>>

\noindent\textbf{Perceptual Performance.} In the original position-length experiment, Cleveland and McGill report that types 1-5 were post-ordered by their perceptual difficulty with type 1 being the easiest and type 5 the hardest. We report the average MLAE for each type across our networks (and total 56 runs per type) as follows: 
for type 1 $MLAE= 3.956 $ ($SD= 0.274 $),
for type 2 $MLAE= 3.952 $ ($SD= 0.441 $),
for type 3  $MLAE= 4.349 $ ($SD= 0.367 $),
for type 4 $MLAE= 3.668 $ ($SD= 0.256 $), and 
for type 5 $MLAE= 3.902 $ ($SD= 0.253 $). These distributions yield significance ($F_{4,25}=2.815, p<0.05$) but post hoc comparions show that really only type 3 and type 4 differ ($t_{10}=3.406, p<0.01$). This means that the networks do not prefer a certain type on average which leads us to \textbf{partially accept H3.1} and we do not replicate the same pattern as Cleveland and McGill in their human studies even though our networks' performance is clearly worse than their human baseline.
\\~\\
\noindent\textbf{`Multi' Network Performance.} In Cleveland and McGill's original position-length experiment, humans were asked to judge visualizations of types 1-5. In the bottom of Figure~\ref{fig:figure4_mlae} we average the human performances of types 1-5 to create a `multi' human. Similarly, when training our networks, we first limit each training to one stimuli. However, a `multi' network which learns types 1-5 simultaenously is closer to the human in the original user study. WHAT SHALL WE TEST HERE?

### 2018-03-31 PAPER/06_bars_and_framed_rectangles_experiment.tex
Minor fixes
Update supplemental.

\section{Bars and Framed Rectangles Experiment}

@@

%\begin{figure}[t]
%	  \includegraphics[width=\linewidth]{figure12_overview}
%  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales. For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles (right). In this case, the right bar is heigher as visible with less free space when adding the frame. We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
%	\label{fig:bars_and_framed_rectangles_experiment}
%\end{figure}
\begin{table}[h]
\centering
\caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales. For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles. In this case, the right bar is heigher as visible with less free space when adding the frame. We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
\resizebox{\linewidth}{!}{
\begin{tabular}{lllr}
	\toprule
	\multicolumn{2}{l}{~} & ~ & Permutations\\

	\midrule
	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure12_0.pdf}} & \makecell[tl]{\emph{Bars}\\~~~Perceptual Task: \emph{Length}\\~~~JND: $X\%$\\~ \\~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $57,600$}\\

	\midrule
	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure12_1.pdf}} & \makecell[tl]{\emph{Framed Rectangles}\\~~~Perceptual Task: \emph{Position}\\~~~JND: $X\%$\\~ \\~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $57,600$}\\

	\bottomrule
\end{tabular}
}
\label{tab:bars_and_framed_rectangles_parameters}
\end{table}

@@

%\begin{figure}[t]
%	  \includegraphics[width=\linewidth]{weber_overview}
%  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution. The lower square contains 10 more dots than the upper one on both sides. However, the difference is easily perceivable on the left while the squares on the right almost look the same. We generate rasterized visualizations similar to this setup and evaluate our classifiers.}
%	\label{fig:webers_law}
%\end{figure}
\begin{table}[h]
\centering
\caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution. We create three different types of images initialized with 10, 100, and 1000 dots. We then mark randomly up to 10  dots in previously free pixels (here visualized in red). For humans, the difference is easily perceivable when the initial dot count is 10 but is hard when it is higher. We evaluate our classifiers on these rasterized visualizations.}
\resizebox{\linewidth}{!}{
\begin{tabular}{lllr}
	\toprule
	\multicolumn{2}{l}{~} & ~ & Permutations\\

	\midrule
	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_0.pdf}} & \makecell[tl]{\emph{Base 10}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\

	\midrule
	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_1.pdf}} & \makecell[tl]{\emph{Base 100}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\

	\midrule
	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_2.pdf}} & \makecell[tl]{\emph{Base 1000}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\

	\bottomrule
\end{tabular}
}
\label{tab:weber_law}
\end{table}

@@

\begin{figure*}[t]
	  \includegraphics[width=\linewidth]{figure12_mlae.pdf}
  \caption{\textbf{Computational results of the Bars-and-Framed-Rectangles experiment.} Log absolute error means and 95\% confidence intervals for the \emph{bars-and-framed-rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}. We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
	\label{fig:figure12_mlae}
\end{figure*}

>>>>

\section{Experiment: Bars and Framed Rectangles}

@@

%\begin{figure}[t]
%	  \includegraphics[width=\linewidth]{figure12_overview}
%  \caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales. For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles (right). In this case, the right bar is heigher as visible with less free space when adding the frame. We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
%	\label{fig:bars_and_framed_rectangles_experiment}
%\end{figure}
%\begin{table}[h]
%\centering
%\caption{\textbf{Bars and Framed Rectangles Experiment.} Cleveland and McGill introduce the bars and framed rectangles experiment which measures the perceptual task of judging position along non-aligned scales. For humans, it is easier to decide which of two bars represent a larger height if a scale is introduced by adding framed rectangles. In this case, the right bar is heigher as visible with less free space when adding the frame. We evaluate whether such a visual aid also helps machines to perceive visually encoded quantities.}
%\resizebox{\linewidth}{!}{
%\begin{tabular}{lllr}
%	\toprule
%	\multicolumn{2}{l}{~} & ~ & Permutations\\
%
%	\midrule
%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure12_0.pdf}} & \makecell[tl]{\emph{Bars}\\~~~Perceptual Task: \emph{Length}\\~~~JND: $X\%$\\~ \\~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $57,600$}\\
%
%
%	\midrule
%	\raisebox{-.85\height}{\includegraphics[width=.5in]{figure12_1.pdf}} & \makecell[tl]{\emph{Framed Rectangles}\\~~~Perceptual Task: \emph{Position}\\~~~JND: $X\%$\\~ \\~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $57,600$}\\
%
%
%	\bottomrule
%\end{tabular}
%}
%\label{tab:bars_and_framed_rectangles_parameters}
%\end{table}

@@

%\begin{figure}[t]
%	  \includegraphics[width=\linewidth]{weber_overview}
%  \caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution. The lower square contains 10 more dots than the upper one on both sides. However, the difference is easily perceivable on the left while the squares on the right almost look the same. We generate rasterized visualizations similar to this setup and evaluate our classifiers.}
%	\label{fig:webers_law}
%\end{figure}
%\begin{table}[h]
%\centering
%\caption{\textbf{Weber-Fechner Law.} The Weber-Fechner law states that the perceivable differences within a distribution is proportional to the initial size of the distribution. We create three different types of images initialized with 10, 100, and 1000 dots. We then mark randomly up to 10  dots in previously free pixels (here visualized in red). For humans, the difference is easily perceivable when the initial dot count is 10 but is hard when it is higher. We evaluate our classifiers on these rasterized visualizations.}
%\resizebox{\linewidth}{!}{
%\begin{tabular}{lllr}
%	\toprule
%	\multicolumn{2}{l}{~} & ~ & Permutations\\
%
%	\midrule
%	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_0.pdf}} & \makecell[tl]{\emph{Base 10}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
%
%
%	\midrule
%	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_1.pdf}} & \makecell[tl]{\emph{Base 100}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
%	
%	\midrule
%	\raisebox{-.85\height}{\includegraphics[width=.5in]{weber_2.pdf}} & \makecell[tl]{\emph{Base 1000}\\~~~JND: $X\%$\\~ \\ ~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~ \\} &~& \makecell[tr]{~\\ $10,000$}\\
%
%	\bottomrule
%\end{tabular}
%}
%\label{tab:weber_law}
%\end{table}

@@

\begin{figure}[t]
	  \includegraphics[width=\linewidth]{figure12_mlae_better_all.pdf}
  \caption{\textbf{Computational results of the Bars-and-Framed-Rectangles experiment.} Log absolute error means and 95\% confidence intervals for the \emph{bars-and-framed-rectangles experiment} as described by Cleveland and McGill~\cite{cleveland_mcgill}. We test the performance of a Multi-layer Perceptron (MLP), the LeNet Convolutional Neural Network, as well as feature generation using the VGG19 and Xception networks trained on ImageNet.}
	\label{fig:figure12_mlae}
\end{figure}

### 2018-03-31 PAPER/SUPPLEMENTAL/supplemental.tex
Update supplemental.

%% indicate IEEE Member or Student Member in form indicated below
%\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
\author{Daniel Haehn, James Tompkin, and Hanspeter Pfister}
\authorfooter{
%% insert punctuation at end of each item
\item Daniel Haehn, and Hanspeter Pfister are with the Paulson School of Engineering and Applied Sciences at Harvard University. \\
E-mail: \{haehn,pfister\}@seas.harvard.edu.
%
\item James Tompkin is with Brown University. \\E-mail: james\_tompkin@brown.edu.
}

@@

\begin{figure}[h]
\centering
\includegraphics[width=\linewidth]{../gfx/figure3_mlae_better_all_AA.pdf}
\caption{\textbf{Anti-aliasing.} We test whether anti-aliasing effects the performance of our networks on pie charts by measuring MLAE. The difference is not statistically significant ($F(1,30)=0.341,p>0.5$).}
\label{fig:aa}
\end{figure}

@@

>>>>

%% indicate IEEE Member or Student Member in form indicated below
%\author{Daniel Haehn, \textit{Member, IEEE}, James Tompkin, and Hanspeter Pfister}
\author{Daniel Haehn, James Tompkin, and Hanspeter Pfister}
\authorfooter{
%% insert punctuation at end of each item
\item Daniel Haehn, and Hanspeter Pfister are with Harvard University. \\
E-mail: \{haehn,pfister\}@seas.harvard.edu.
%
\item James Tompkin is with Brown University. \\E-mail: james\_tompkin@brown.edu.
}

@@

\begin{figure}[h]
\centering
\includegraphics[width=\linewidth]{../gfx/figure3_mlae_better_all_AA.pdf}
\caption{\textbf{Anti-aliasing.} We test whether anti-aliasing effects the performance of our networks on pie charts by measuring MLAE. The difference is not statistically significant ($F(1,30)=0.341,p>0.5$).}
\label{fig:aa}
\end{figure}
\vfill\null
\section{Experiment: Noise}
For all our experiments, we add subtle $5\%$ noise to every pixel to enhance variability. We did not observe a significant effect on regression performance when comparing the weber-fechner's law experiment with and without noise averaged over 4 runs (Figure~\ref{fig:noise}). However, the variability of our stimuli increases.

\begin{figure}[th]
\centering
\includegraphics[width=.8\linewidth]{../gfx/weber_mlae_no_noise.pdf}
\includegraphics[width=.8\linewidth]{../gfx/weber_mlae_noise.pdf}
\caption{\textbf{Noise.} We test whether noise (top: off, bottom: on) effects the performance of our networks on the weber-fechner's law experiment by measuring MLAE. With noise, $MLAE=4.511$ ($SD=0.512$) and without noise $MLAE=4.491$ ($SD=0.543$). The difference is not statistically significant ($F(1,22)=0.008,p>0.5$).}
\label{fig:noise}
\end{figure}

@@

\begin{figure}[p]
	\centering
	  \includegraphics[width=\linewidth]{../gfx/figure4_training_loss_with_multi.pdf}
	  \includegraphics[width=\linewidth]{../gfx/figure4_val_loss_with_multi.pdf}
  \caption{\textbf{Loss plots for the position-length experiment.} We visualize the MSE loss on training data and for unseen validation data after each epoch. There is no significant difference in convergence for either encoding but spiky outliers due to monte-carlo cross validation are visible.}
	\label{fig:fig4_loss}
\end{figure}

\begin{figure}[p]
	\centering
	  \includegraphics[width=\linewidth]{../gfx/figure12_training_loss.pdf}
	  \includegraphics[width=\linewidth]{../gfx/figure12_val_loss.pdf}
  \caption{\textbf{Loss plots for the bars-and-framed-rectangles experiment.} We visualize the MSE loss on training data and for unseen validation data after each epoch. There is no significant difference in convergence for either encoding.}
	\label{fig:fig12_loss}
\end{figure}

\begin{figure*}[t]
	\centering

    \subfloat[without noise]{
	  \includegraphics[width=.48\linewidth]{../gfx/weber_training_loss_no_noise.pdf}
	  \includegraphics[width=.48\linewidth]{../gfx/weber_validation_loss_no_noise.pdf}
	}
	\hfill
    \subfloat[with noise]{
	  \includegraphics[width=.48\linewidth]{../gfx/weber_training_loss.pdf}
	  \includegraphics[width=.48\linewidth]{../gfx/weber_validation_loss.pdf}
	}

  \caption{\textbf{Loss plots for the weber-fechner's law experiment.} We visualize the MSE loss on training data (left) and for unseen validation data (right) after each epoch (a) without noise and (b) with subtle $5\%$ noise per pixel. There is no significant difference when noise is added. The LeNet network seems to overfit with Weber Base 100 in both cases even with dropout regularization.}
	\label{fig:weber_loss}
\end{figure*}
